A boundary-integral framework to simulate \\ the viscous erosion of a porous medium

### 2018-04-06 docs/methods/erosion_methods.tex
small changes

Recent work in the high-Reynolds number regime has highlighted the
importance of the feedback between flow and morphology that occurs
during erosion~\cite{ris-moo-chi-she-zha2012, moo-ris-chi-zha-she2013,
hewett2017evolution, moore2017riemann}, and a similar feedback occurs in
the related processes of dissolution and melting~\cite{Huang2015,
kondratiuk2015steadily, rycroft2016asymmetric, cohen2016erosion,
hewett2017pear, claudin2017dissolution}. In such cases, the flow alters
the shape of the immersed structures, which in turn modifies the
surrounding flow, and so on. Collectively, such processes have been
referred to as `sculpting with flow'~\cite{ristroph2018sculpting}. From
this viewpoint, the example of an eroding porous medium presents new
opportunities and challenges. First, the medium itself is composed of
many solid constituents, so multi-body interactions, especially close
interactions, must be considered. Second, the long-range interactions
facilitated by Stokes flow makes all the interactions between bodies important.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
To perform this simulation of Stokesian erosion of a porous media, two
fast and accurate methods are employed---one to solve the fluid
equations and the other to integrate the eroding grains.  The fluid
equations are reformulated in terms of a boundary integral equation
(BIE).  Since the boundary conditions are Dirichlet, we use a
double-layer potential formulation so that the resulting BIE is
second-kind.  This second-kind BIE is discretized with the spectrally
accurate trapezoid rule~\cite{tre-wei2014} which results in a
well-conditioned dense linear system for an unknown density function.
This linear system is solved using fast-multipole-accelerated
GMRES~\cite{saa-sch1986}.  Once the density function is computed, the
erosion rate (shear stress) is computed with a spectrally accurate
odd-even quadrature rule~\cite{sid-isr1988}.  In addition, the pressure
and vorticity are computed in the fluid bulk (Fig.~\ref{fig:50bodies})
with an upsampled trapezoid rule in the fluid bulk to assist with
visualizing the effect of erosion in porous media.

@@

To simulate multiple bodies, it is essential to resolve the nonlocal
effects between the different bodies.  In this work, these interactions
are included with high-order accuracy by using an integral equation
method~\cite{poz1992}.  Integral equations have been used extensively to
simulate viscous fluids in complex geometries, including porous media
flow~\cite{dea-qua-bir-jua2018, bar-mar-vee-zha2016}. Challenges of a
BIE formulation include resolving the interactions between nearly
touching bodies and developing efficient preconditioners for the
discretized system.  Possible preconditioners include
multigrid~\cite{qua-bir2015a} and the inverse fast multipole
method~\cite{qua-cou-dar2018}, but the results presented in this work
require no more than $\mathcal{O}(100)$ GMRES iterations which is too
small to make preconditioners for BIEs effective; preconditioners will
be considered further in future work when there are many more grains
that are also being transported by the flow.  To address the
nearly-singular integrands, we use the fact that the grains are eroding
and not moving, so their separation increases with time.  Therefore, in
this work, we apply an upsampling scheme, but in future work, we will
require a near-singular integration scheme~\cite{qua-bir2014a,
klo-bar-gre-one2013, bar-wu-vee2015, hel-oja2008a, bea-lai2001}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions\label{s:conclusions}}

>>>>

Recent work in the high-Reynolds number regime has highlighted the
importance of the feedback between flow and morphology that occurs
during erosion~\cite{ris-moo-chi-she-zha2012, moo-ris-chi-zha-she2013,
hewett2017evolution, moore2017riemann}, and a similar feedback occurs in
the related processes of dissolution and melting~\cite{Huang2015,
kondratiuk2015steadily, rycroft2016asymmetric, cohen2016erosion,
hewett2017pear, claudin2017dissolution}. In such cases, the flow alters
the shape of the immersed structures, which in turn modifies the
surrounding flow, and so on. Collectively, such processes have been
referred to as `sculpting with flow'~\cite{ristroph2018sculpting}. From
this viewpoint, the example of an eroding porous medium presents new
opportunities and challenges. For example, the medium itself is composed
of many solid constituents, so multi-body interactions, especially close
interactions, must be considered.
%Second, the long-range interactions
%facilitated by Stokes flow makes all the interactions between bodies important.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
To perform this simulation of Stokesian erosion of a porous media, two
fast and accurate methods are employed---one to solve the fluid
equations and the other to integrate the eroding grains.  The fluid
equations are reformulated in terms of a boundary integral equation
(BIE)~\cite{poz1992}.  Since the boundary conditions are Dirichlet, we
use a double-layer potential formulation so that the resulting BIE is
second-kind.  This second-kind BIE is discretized with the spectrally
accurate trapezoid rule~\cite{tre-wei2014} which results in a
well-conditioned dense linear system for an unknown density function.
This linear system is solved using fast-multipole-accelerated
GMRES~\cite{saa-sch1986}.  Once the density function is computed, the
erosion rate (shear stress) is computed with a spectrally accurate
odd-even quadrature rule~\cite{sid-isr1988}.  In addition, the pressure
and vorticity are computed in the fluid bulk (Fig.~\ref{fig:50bodies})
with an upsampled trapezoid rule in the fluid bulk to assist with
visualizing the effect of erosion in porous media.

@@

To simulate Stokesian erosion, it is essential to resolve the nonlocal
effects between the different bodies.  BIEs naturally resolve these
interactions, and they have been used extensively to simulate viscous
fluids in complex geometries, including porous media
flow~\cite{dea-qua-bir-jua2018, bar-mar-vee-zha2016}. Challenges of a
BIE formulation include resolving the interactions between nearly
touching bodies and developing efficient preconditioners for the
discretized system.  Possible preconditioners include
multigrid~\cite{qua-bir2015a} and the inverse fast multipole
method~\cite{qua-cou-dar2018, cou-pou-dar2017}, but the results
presented in this work require no more than $\mathcal{O}(100)$ GMRES
iterations which is too small to make preconditioners for BIEs
effective; preconditioners will be considered in future work when many
more grains are eroding and being transported by the flow.  To address
the nearly-singular integrands, we use the fact that the grains are
eroding and not moving, so their separation increases with time.
Therefore, in this work, we apply an upsampling scheme, but in future
work, we will require a near-singular integration
scheme~\cite{qua-bir2014a, klo-bar-gre-one2013, bar-wu-vee2015,
hel-oja2008a, bea-lai2001}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions\label{s:conclusions}}
We have developed a numerical framework for simulating two-dimensional
Stokesian erosion in a porous media.  By using a BIE formulation of the
fluid equations and a {\thL}, we achieve spectral accuracy
